---
title: "STAT 423 Project"
output: html_document
date: "2025-04-02"
---

```{r setup, include=FALSE}
install.packages("mosaic")
library("mosaic")
install.packages("ggplot2")
library(ggplot2)


install.packages("DescTools")
library(DescTools)

Skip_class <- read.csv("https://raw.githubusercontent.com/badongskii/stat423_skipping_class_survey/refs/heads/main/skippingclass_cleaned.csv")
attach(Skip_class)
Skip_class
names(Skip_class)

```
Applying modal imputation to fill non-responses
```{r}
#summary to show quick stats of each column in csv file. Reveals how many non responses in each.
summary(Skip_class)

```
Code to check which indices are n/a
```{r}
#after seeing which columns had non responses, code below indicates which indicies have the non responses
which(is.na(Skip_class$Percentage.Skip))
which(is.na(Skip_class$Types.Classes))
which(is.na(Skip_class$Options.Main))
which(is.na(Skip_class$Review.missed.material))
```
Missing data in Percentage.Skip, Types.Classes, Options.Main and Review.missed.material.
Find mean for each column which will be imputed later.
Note: decided to do mode since almost all questions are categorical.
```{r}
mean.percentage.skip <- favstats(~Percentage.Skip, data = Skip_class)$mean


Skip_class.imputed <- Skip_class #copy of the data set

Skip_class.imputed$Percentage.Skip[which(is.na(Skip_class$Percentage.Skip))] <- mean.percentage.skip


#the mode function 
#Mode(Skip_class$Percentage.Skip, na.rm = T)
#mode.percentage.skip <- Mode(Skip_class$Percentage.Skip, na.rm = T)[[1]]
#Skip_class.imputed$Percentage.Skip[which(is.na(Skip_class$Percentage.Skip))] <- mode.percentage.skip

Mode(Skip_class$Types.Classes, na.rm = T) #computes the mode of the types of classes before imputing the data 
mode.type.classes <- Mode(Skip_class$Types.Classes, na.rm = T)[[1]] #this chooses the mode
Skip_class.imputed$Types.Classes[which(is.na(Skip_class$Types.Classes))] <- mode.type.classes #imputes the mode into the n/a within the types.classes data set

Mode(Skip_class$Options.Main, na.rm = T) #computes the mode of options.main before imputing the data
mode.options.main <- Mode(Skip_class$Options.Main, na.rm = T)[[1]] #this chooses the mode to impute
Skip_class.imputed$Options.Main[which(is.na(Skip_class$Options.Main))] <- mode.options.main #imputes the data into the n/a within the options.main

Mode(Skip_class$Review.missed.material, na.rm = T) #computes the mode of Review.missed.material before imputing the data
mode.review.material <- Mode(Skip_class$Review.missed.material, na.rm = T)[[1]] #this chooses the mode to impute
Skip_class.imputed$Review.missed.material[which(is.na(Review.missed.material))] <- mode.review.material #imputes the data into the n/a within the Review.missed.material

Skip_class.imputed
```


```{r}
summary(Skip_class.imputed)
```
Data is now cleaned.


#Compute a linear regression analyisis to determine our most important variables:
```{r}
skip_mod1 <- lm(`Percentage.Skip`~Gender+`Year.Study`+Courses+Reason+`Types.Classes`+`Options.Main`+`Review.missed.material`+`Grades.affected`+`Transport`+`Factor.not.to.skip`, data = Skip_class.imputed)
summary(skip_mod1)

```


## Get the mean values from the data collected:


```{r}
favstats(~Reason, data = Skip_class.imputed)
favstats(~Year.Study, data = Skip_class.imputed)
favstats(~Options.Main, data = Skip_class.imputed)
favstats(~Factor.not.to.skip, data = Skip_class.imputed)


```
$$
\begin{array}{lccc}
                           &  Mean   &  Sd    &  n       \\[0.5ex]
\hline
\text{Reason}    &   4.67     &   2.98      & 40                     \\[0.5ex]
\text{Year of Study}          &   2.5      &   1.13       &  40    &       \\[0.5ex]
\text{Options or Main Classes} &  1.27  & 0.452    & 40     \\[0.5ex]  
\text{Factor to not Skip Class}   &  2.7 & 1.02 & 40     \\[0.5ex]

\hline 
\end{array}
$$

## Including Plots


```{r}
hist(`Reason`,col="tomato2")
hist(`Year.Study`,col="gold2")
hist(`Options.Main`, col="lightskyblue")
hist(`Factor.not.to.skip`, col="lightcyan3")
```
#Reasons students would skip class
```{r}
Science_pop <- 7400
Sample_sn <- 40
Study_num <- 14
Not_useful <- 13
Other_reason <- 7
Study_p <- 14/40
NUsef_p <- 13/40
Other_p <- 7/40
Person_p <- 2/40

moe_stud <- qnorm(.975)*sqrt((Study_p*(1-Study_p)/39)*((Science_pop-Sample_sn)/Science_pop))
moe_stud
lowstudy <- Study_p-moe_stud
upstudy <- Study_p+moe_stud
lowstudy
upstudy

moe_usef <- qnorm(.975)*sqrt((NUsef_p*(1-NUsef_p)/39)*((Science_pop-Sample_sn)/Science_pop))
moe_usef
lowusef <- NUsef_p-moe_usef
upusef <- NUsef_p+moe_usef
lowusef
upusef

moe_other <- qnorm(.975)*sqrt((Other_p*(1-Other_p)/39)*((Science_pop-Sample_sn)/Science_pop))
moe_other
lowother <- Other_p-moe_other
upother <- Other_p+moe_other
lowother
upother

moe_person <- qnorm(.975)*sqrt((Person_p*(1-Person_p)/39)*((Science_pop-Sample_sn)/Science_pop))
moe_person
lowperson <- Person_p-moe_person
uperson <- Person_p+moe_person
lowperson
uperson

misc.p <- 6/40

moe_misc <- qnorm(0.975)*sqrt((misc.p*(1-misc.p)/39)*((Science_pop-Sample_sn)/Science_pop))

misc.lb <- misc.p - moe_misc
misc.ub <- misc.p + moe_misc
```

Data visualization
```{r}
#creating a vector with all possible responses for reasons students would not skip class
responses <- c("Study", "Not Interesting/Useful", "Other", "3 Misc. Options")

#combined all proportions, lower bounds and upper bounds of each response and saved them in corresponding vectors
prop <- c(Study_p, NUsef_p, Other_p, misc.p)
lb <- c(lowstudy, lowusef, lowother, misc.lb)
ub <- c(upstudy, upusef, upother, misc.ub)


data <- data.frame(responses, prop, lb, ub)

#plotting all the confidence intervals and answer proportions into a bar plot to compare each response
ggplot(data, aes(x = responses, y = prop, fill = responses)) + 
  geom_bar(stat = "identity") +
  geom_errorbar(
    aes(ymin =lb, ymax =ub),  
    color = "black") + 
    
  labs(title = "Reasons for Students Not to Skip Class (Proportion with 95% CI)", x = "Response", y = "Proportion") +
  scale_color_brewer(palette = "Paired")


```


##Reasons not to skip class (Population Proportion Estimates)

Reason to not skip class proportion (Top 2)
```{r}
n <- 40 #sample size
N <- 7400 #population size

most.combined <- sum(Skip_class.imputed$Factor.not.to.skip == 2) + 
                  sum(Skip_class.imputed$Factor.not.to.skip == 3) #combining the count of the top two answers chosen into a single variable

prop.most.combined <- most.combined / n #sample proportion of the top two most chosen response

#CI with 95% confidence
z <- qnorm(0.975)
most.combined.error <- sqrt((prop.most.combined*(1 - prop.most.combined)/(n - 1))*((N - n)/N)) #error calculation

most.combined.lb <- prop.most.combined - z * most.combined.error #lower bound for CI
most.combined.ub <- prop.most.combined + z * most.combined.error #upper bound for CI

cat("95% CI: [", most.combined.lb, ",", most.combined.ub, "]\n")
```
Conclusion: We can say with 95% confidence that between 50% and 80% of science students would choose not to skip class due to required attendance and lack of online material.

BOOTSTRAPPING TOP 2 REASONS NOT TO SKIP
```{r}
nsize <- 40

most.combined <- sum(Skip_class.imputed$Factor.not.to.skip == 2) + 
                  sum(Skip_class.imputed$Factor.not.to.skip == 3)

prop.most.combined <- most.combined / n
prop.most.combined

notconfident.vector <- c(rep(0, nsize*(1-prop.most.combined)))
confident.vector <- c(rep(1, nsize*(prop.most.combined)))

main.not.skip <- c(notconfident.vector, confident.vector)

Nbootstrap <- 5000

samp.bootstrap <- numeric(Nbootstrap)

for(i in 1:Nbootstrap)
{numsuccess <- sum(sample(main.not.skip, nsize, replace = T))
 samp.bootstrap[i] <- (numsuccess) / nsize
}

prop.bootstrap1 <- data.frame(samp.bootstrap)

```
Checking the prop.bootstrap1 data frame
```{r}
head(prop.bootstrap1, 5)
```

```{r}
tail(prop.bootstrap1, 5)
```
Visualization of the bootstrap proportion
```{r}
ggplot(data=prop.bootstrap1, aes(x = samp.bootstrap)) + geom_histogram(col='black', fill='lightblue', binwidth=0.025) + xlab("Values of Bootstrap Statistic (Sample Proportion)") + ylab("Count") + ggtitle("Bootstrap Distribution of Sample Proportion")

```
Calculating the bootstrap proportion CI with 95% confidence
```{r}
qdata(~samp.bootstrap, c(0.025, 0.975), prop.bootstrap1)
```
Results are basically identical to original CI calculation


Finding the proportion for each option on reasons science students attend class. 
First response was commented. Same procedure for the other questions.
```{r}
participation <- sum(Skip_class.imputed$Factor.not.to.skip == 1) #counts the number of participation responses in the csv file
participation.p <- participation / n #find the sample proportion by dividing by n=40

participation.error <- z * sqrt((participation.p*(1 - participation.p)/(n - 1))*((N - n)/N)) #error calculation 

participation.lb <- participation.p - participation.error #calculated lower bound for CI
participation.ub <- participation.p + participation.error #calculated upper bound for CI

c(participation.lb, participation.ub)
```

```{r}
marks <- sum(Skip_class.imputed$Factor.not.to.skip == 2)
marks.p <- marks / n

marks.error <- z * sqrt((marks.p*(1 - marks.p)/(n - 1))*((N - n)/N))

marks.lb <- marks.p - marks.error
marks.ub <- marks.p + marks.error

c(marks.lb, marks.ub)
```

```{r}
no.material <- sum(Skip_class.imputed$Factor.not.to.skip == 3)
no.material.p <- no.material / n

no.material.error <- z * sqrt((no.material.p*(1 - no.material.p)/(n - 1))*((N - n)/N))

no.material.lb <- no.material.p - no.material.error
no.material.ub <- no.material.p + no.material.error

c(no.material.lb, no.material.ub)
```

```{r}
being.diligent <- sum(Skip_class.imputed$Factor.not.to.skip == 4)
being.diligent.p <- being.diligent / n

being.diligent.error <- z * sqrt((being.diligent.p*(1 - being.diligent.p)/(n - 1))*((N - n)/N))

diligent.lb <- being.diligent.p - being.diligent.error
diligent.ub <- being.diligent.p + being.diligent.error

c(diligent.lb, diligent.ub)
```

#Barplot visualization for reasons not to skip class
```{R}
#creating a vector with all possible responses for reasons students would not skip class
possible.responses <- c("Participation", "Marks", "Lack of online material", "Being diligent")

#combined all proportions, lower bounds and upper bounds of each response and saved them in corresponding vectors
all.prop <- c(participation.p, marks.p, no.material.p, being.diligent.p)
all.lb <- c(participation.lb, marks.lb, no.material.lb, diligent.lb)
all.ub <- c(participation.ub, marks.ub, no.material.ub, diligent.ub)

plot_data <- data.frame(possible.responses, all.prop, all.lb, all.ub)

#plotting all the confidence intervals and answer proportions into a bar plot to compare each response
ggplot(plot_data, aes(x = possible.responses, y = all.prop, fill = possible.responses)) + 
  geom_bar(stat = "identity") +
  geom_errorbar(
    aes(ymin = all.lb, ymax = all.ub),  
    color = "black") + 
    
  labs(title = "Reasons for Students Not to Skip Class (Proportion with 95% CI)", x = "Response", y = "Proportion") +
  scale_color_brewer(palette = "Paired")

```



Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
```{r}
#Mean and Confidence Intervals for Percentage Classes Skipped

n<-nrow(Skip_class)
#total amount of science undergraduates
N<-7400

#Sample mean
ybar<-mean(Skip_class.imputed$Percentage.Skip)

#Sample sd
s<- sd(Skip_class.imputed$Percentage.Skip)

z<-qnorm(0.975)

#Confidence interval
lowbo<-ybar-z*sqrt(1-(n/N))*(s/sqrt(n))
upbo<-ybar+z*sqrt(1-(n/N))*(s/sqrt(n))

ybar
c(lowbo, upbo)
```


Attempting bootstrap with skip class percentages
```{r}
Nresamples <- 5000 #sets the number of re samples to 5000
n <- length(Skip_class.imputed$Percentage.Skip) #receives the sample size of the original sample survey which is 40

bootmean.skip <- numeric(Nresamples) #creates a vector that holds all the 5000 values of the bootstrap statistic which is the mean here

bootsd.skip <- numeric(Nresamples) #does the same as the previous vector but for standard deviation

for(i in 1:Nresamples) #initiates for loop
{ bootdata <- sample(Skip_class.imputed$Percentage.Skip, n, replace = T) #resamples the 40 percentages then stored in bootdata

  bootmean.skip[i] <- mean(bootdata) #computes the mean of each data 
  bootsd.skip[i] <- sd(bootdata)
}

bootstrap.skip.df = data.frame(bootmean.skip, bootsd.skip)

```

```{r}

tail(bootstrap.skip.df) #checking the data

```

```{r}

favstats(bootmean.skip, bootstrap.skip.df)

```
Visualizations of bootstrap

```{r}

ggplot(data=bootstrap.skip.df, aes(x = bootmean.skip)) + geom_histogram(col='black', fill='lightblue', binwidth=1) + xlab("Values of Bootstrap Statistic (Mean)") + ylab("Count") + ggtitle("Bootstrap Distribution of the Sample Mean")

```

```{r}

qdata(~bootmean.skip, c(0.025, 0.975), data = bootstrap.skip.df) #mean confidence interval of Percentage of Skipping class (95% confidence)

```
We can infer that bootstrapping in this case does not show a noticeable difference when calculating the CI of skipping class percentage.


```{r}
# Define variables
y<-Skip_class.imputed$Percentage.Skip
x <- Skip_class.imputed$Year.Study

# Sample size and population size

# Sample means
xbar <- mean(x)

# Known population mean of x (Year of Study)
# If unknown, use xbar as estimate
xU <- 2.5  # Given or assumed known

# Regression coefficient (slope)
b1 <- sum((x - xbar) * (y - ybar)) / sum((x - xbar)^2)

# Regression estimate of population mean
yreg <- ybar + b1 * (xU - xbar)

# Standard error for regression estimator (with FPC)
residuals <- y - (b1 * x + (ybar - b1 * xbar))  # regression residuals
s2e <- sum(residuals^2) / (n - 2)
seyreg <- sqrt((1 - n/N) * (s2e / n))

# 95% Confidence Interval
z <- qnorm(0.975)
reglower <- yreg - z * seyreg
regupper <- yreg + z * seyreg

# Output results
yreg
c(reglower, regupper)
```





##Boxplot visualization for the class skip percentage
```{r}
ggplot(Skip_class.imputed, aes(y = Percentage.Skip)) +
  geom_boxplot(fill = "lightblue", color = "black") +
  labs(title = "Distribution of Percentage of Classes Skipped", y = "Percentage Skipped (%)")

```


##Boxplot visualization for the class skip percentage in correlation to year of study
```{r}

ggplot(Skip_class.imputed, aes(x = Year.Study, y = Percentage.Skip, fill = Year.Study)) +
  geom_boxplot() +
  labs(title = "Percentage Skipped by Year of Study",
       x = "Year of Study",
       y = "Percentage Skipped in %") +
  scale_color_brewer(palette = "pastell1")


```



```{r}
# Ratio estimate of yU
ybar <- mean(y)
xbar <- mean(x)
xU <- 2.5  # use known population mean if known, or sample mean
ratioestimate <- (ybar / xbar) * xU
```