---
title: "STAT 423 Project"
output: html_document
date: "2025-04-02"
---

```{r setup, include=FALSE}
install.packages("mosaic")
library("mosaic")
install.packages("ggplot2")
library(ggplot2)
install.packages("DescTools")
library(DescTools)

Skip_class <- read.csv("https://raw.githubusercontent.com/badongskii/stat423_skipping_class_survey/refs/heads/main/skippingclass_cleaned.csv")
attach(Skip_class)
Skip_class
names(Skip_class)
```

Applying modal imputation to fill non-responses
```{r}
#summary to show quick stats of each column in csv file. Reveals how many non responses in each.
summary(Skip_class)

```
Code to check which indices are n/a
```{r}
which(is.na(Skip_class$Percentage.Skip))
which(is.na(Skip_class$Types.Classes))
which(is.na(Skip_class$Options.Main))
which(is.na(Skip_class$Review.missed.material))
```
Missing data in Percentage.Skip, Types.Classes, Options.Main and Review.missed.material.
Find mean for each column which will be imputed later.
Note: decided to do mode since almost all questions are categorical.
```{r}
#mean.percentage.skip <- favstats(~Percentage.Skip, data = Skip_class)$mean
#mean.types.classes <- favstats(~Types.Classes, data = Skip_class)$mean
#mean.options.main <- favstats(~Options.Main, data = Skip_class)$mean
#mean.review.material <- favstats(~Review.missed.material, data = Skip_class)$mean

Skip_class.imputed <- Skip_class #copy of the data set

#Skip_class.imputed$Percentage.Skip[which(is.na(Skip_class$Percentage.Skip))] <- mean.percentage.skip
#Skip_class.imputed$Types.Classes[which(is.na(Skip_class$Types.Classes))] <- mean.types.classes
#Skip_class.imputed$Options.Main[which(is.na(Skip_class$Options.Main))] <- mean.options.main
#Skip_class.imputed$Review.missed.material[which(is.na(Review.missed.material))] <- mean.review.material

Mode(Skip_class$Percentage.Skip, na.rm = T)
mode.percentage.skip <- Mode(Skip_class$Percentage.Skip, na.rm = T)[[1]]
Skip_class.imputed$Percentage.Skip[which(is.na(Skip_class$Percentage.Skip))] <- mode.percentage.skip

Mode(Skip_class$Types.Classes, na.rm = T)
mode.type.classes <- Mode(Skip_class$Types.Classes, na.rm = T)[[1]]
Skip_class.imputed$Types.Classes[which(is.na(Skip_class$Types.Classes))] <- mode.type.classes

Mode(Skip_class$Options.Main, na.rm = T)
mode.options.main <- Mode(Skip_class$Options.Main, na.rm = T)[[1]]
Skip_class.imputed$Options.Main[which(is.na(Skip_class$Options.Main))] <- mode.options.main

Mode(Skip_class$Review.missed.material, na.rm = T)
mode.review.material <- Mode(Skip_class$Review.missed.material, na.rm = T)[[1]]
Skip_class.imputed$Review.missed.material[which(is.na(Review.missed.material))] <- mode.review.material

Skip_class.imputed
```


```{r}
summary(Skip_class.imputed)
```
Data is now cleaned.




## Get the mean values from the data collected:
```{r cars}
favstats(~Gender, data = Skip_class.imputed)
favstats(~Gender, data = Skip_class.imputed)
favstats(~Gender, data = Skip_class.imputed)

```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
